---
title: "Slangbos encroachment - Mini Paper"
author: "Konstantin Schellenberg"
date: "`r format(Sys.time(), '%B %e, %Y')`"
output: 
    bookdown::html_document2
bibliography: [slangbos.bib, refs.bib]
biblio-style: apalike
always_allow_html: true
---

<style>
.html-widget {
    margin: auto;
}
</style>

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE, cache=TRUE}

my.packages = c("raster", "sf", "rgdal", "caret")
lapply(my.packages, require, character.only = TRUE)

source("import.R")
```


```{r include=FALSE}
knitr::write_bib(my.packages,
  file = "refs.bib",
  prefix = "R-"
)
knitr::opts_chunk$set(fig.align = "center")
```

# Introduction

The increase of woody cover in the semi-arid and arid environments have especially pulled consequences for the savannas on the African continent [@Eldridge.2011]. @Stevens.2016 have shown that woody cover has increased throughout the South African savannas within a 70 years period. Both, local and global drivers are suspected to drive encroachment. The local extinction of large herbivore herds and the increase of global CO<sub>2</sub> in the atmosphere are likely to be the main indicators for the expansion of woody biomass increase in African savanna. CO<sub>2</sub> fertilisation, meaning the acceleration of plant growth due to elevated total carbon amount in the atmosphere, is said to improve the water-use efficiency as for the reduced stomatal conductance of woody plants [@Ainsworth.2005]. As consequence, higher C0<sub>2</sub> with constant water ingestion lead to higher growth rates. Still, water availability and temperature are key constraints on woody plant growth and are the crucial parameters in explaining shrub encroachment patterns [@Stevens.2016].

Though being native to the South African province of Western Cape the dwarf shrub Seriphium plumosum (Slangbos, Bankrupt Bush) has been documented to be the main encroacher on the grassvelds (South African grassland biomes) in the provinces of Free State, North West, Mpumalanga, Eastern Cape and Gauteng Provinces [@mucina.2006, @morgenthal.2018]. The shurb expects precipitation of about 620–750 mm [@snyman.2009], reflecting the boundary between the semi-arid and the mesic biome. More recently, S. plumosum has been found encroaching into more arid regions [@synman.2009].
@synman.2012 describes the habitat of S. plumosum as wet, with perennial and transient water bodies, relatively high soil fertility and loamy soils. All which is being gound on old, overgrazed and overfertilised rangeland and fields. The species has light-coloured, thus sun-reflecting leaves attached to long and thin branches being aligned vertically and has unique spectral characteristics [@dubula.2016, @tesfamichael.2018]. These features together with oily compartments makes the shrub well adapted to long, hot summers. This makes it eventually more competitive than grass communities originally found on the encroachment plots.

In the case of the grassy plains of Free State slangbos communities increasingly drive out the prevalent open and grassy landscapes. 11 million hectares of rangeland that could be unsuitable for agricultural production [@avenant.2015]. The rapid change of the land's botanic composition pose great challenges for the local ecological dynamics on the one, and for livestock breeders and shepherds on the other hand. 
An overall decrease of phytomass production on encroached sites has a significant relationship to encroachment. Due to the higher ability of Slangbos roots to compete with grass for water [@synman.2012], grassland is supressed and with it, the annual rainfall biomass stock growths.
Also, Common livestock species avoid ingesting the hard and bitter-oily structure of Slangbos. Consequently, being left aside, the shrub imposes higher pressure on the existing open grassland. Remaining grass plots suffer under increasing overgrazing and land degradation while removing it is both, cost and labour intensive. 

Mapping the scope and speed of encroachment patterns is of crucial value for managing this large scale biome change. Remote Sensing time series have proven to be an effective source for ecological change mapping. With growing availability of freely available optical and SAR satellite data analyses are cheap, fast and reproducible. The ESA's Sentinel-1/2 provide freely accessible high resolution time series of weekly (and less) coverage are seen to be the new working horse of land surface interpretation.

Bush encroachment mapping has been tackled by recent studies [@ludwig.2019, @morgenthal.2016, @morgenthal.2018, @marshalane.2018] and is closely related to land cover analyses [@ngi, ...] and woody cover mapping [@urban.2020, ...]. 

Integrated use of radar (Sentinel-1) and optical (Sentinel-2) signal in moderate spatial and high temporal resolution has already returned promising results [@steinhausen.2018, @urban.2018, zhang.2019] resulting in higher accuracies than single with single use. This is due to the different surface properties they are sensitive to. 

This study aims to use primarily SAR (Synthetic Aperture Radar) time series data in order to measure bush encroachment on the both, the spatial and the temporal scale. Five years of Sentinel-1 C-Band data available.
Preprocessing of the Sentinel-1 data has been done by Dr Marcel Urban (Friedrich Schiller University Jena). Hirner (DLR) is responsible for the Sentinel-2 pre-processing chain. 

machine learning, Random forest good for remote sensing [@Ludwig.2019]

## Data exploration

Drought events in summer seasons of 2015 and 2016 due to the exceptionally strong El Niño

## Summary of introduction

Yet, none of the studys have taken coherences into account, neither time series of SAR data.


---
# Study Area and Data

sentinel-2 granule
sentinel-1 product and pre-processing


# Methods

Training sample were retrieved from aerial imagery [@NGI] and Google Earth VHR data [@GoogleEarth] via a visual and manual picking approach. Ground truth information and expert knowledge has been retrieved by survey on site and a photo documentation. 

The random forest algorithm [@breiman.2001] has been successfully used for multi-sensor and spatio-temporal applications [@belgiu.2016, @corcoran.2013]. The model is trained on the relationship of bush structure and backscatter signal to gain a sensitivity for subtle feature discrimination in hypertemporal radar data cubes. More on Sentinel 1...

To enhance the sensitivity of the classification problem, further optical information was incorporated in the model to enhance the variable set. @Ludwig.2019 has shown that the NDVI supplying bands, red and near infrared (NIR) deliver highest accordance with savanna bush image patterns. As for this finding, Sentinel-2's 4th and 8th band time series was used over the entire acquisition time (2016 - ). Cloud masking, gap filling via OTB::ImageTimeSeriesGapFilling[@otbgapfilling.2016]. Gap filling was necessary as classification algorithms do not expect NA information as input dataset. Hence, to hold a large optical data source, gap filling with < 20% cloud covered scenes seems appropriate 

The structural element of slangbos are likely to be recorded by C-Band signals. The fraction of bush compartments are roughly of the size where C-Band waves interact the most. Additionally, the round bush geometry implies it to be a strong volume scatterer. The depolarisation leads to a large cross-pol fraction. First investigations on the reaction of slangbos in C-Band waves implies more VH backscatter with growing bush volume and bush cover density the VH [\@ref(fig:exploration-plot)].

### Characteristics of Seriphium plumosum
 wet conditions, perennial and transient water tables, overly high soil pH, clay content, high soil fertility (organic matter and phosphate) and sodium chloride. The deeper the soil profile, the more favourable the habitat for S. plumosum. [@snyman.2012]
 

[@morgenthal.2018]


```{r exploration-plot, echo=FALSE, fig.cap="Median and standard deviation interval of Sentinel-1 VH time series (temporal resolution less than 7d, 354 scenes). Sites associated with slangbos (Seriphium plumosum) return a stronger signal the further the bush encroached (red, n = 564 pixels). Agricultural sites leave stronger patterns of seasonality without overall slope (blue, n = 1393 pixels). Plots being subject to slangbos control through chemical treatment, burning or manual uprooting or ploughing return significantly less signal after the operation (black, n = 105 pixels)."}
"plt"
```

## notes for methods

- Sentinel-2 scene with >20% NA deleted
- cloud mask applied
- cropped to area extent
- band 4 (red): 
- band 8 (near-infrared): true 10m resolution, bush have large contrast to vegetation
- ndvi

- Intelligent cloud gap filling performed via OTB::ImageTimeSeriesGapFilling[@otbgapfilling.2016]. Here, missing information of a cloud covered scene is been substitute by interpolated image response of the respective neighbor in time, 1 before and 1 after the acquisition. The DINEOF algorithm did not work. Too coarse...

Training a learner means fitting a model to a given data set. Subsequently, we want to predict the label for new observations. These predictions are compared to the ground truth values in order to assess the predictive performance of the model. <https://mlr3book.mlr-org.com/train-predict.html>

# Results

# Discussion

This large error can be attributed to canopy
spacing which, within an image, increases the reflectance of other vegetation and soil in the
neighbourhood of S. plumosum, reducing reflectance purity of the S. plumosum canopy [@Mashanale.2016]

Reflectance:

- Stable NDVI throughout the year. Grass & Forest much more seasonality
- S2 Bands 4, 7 suitable [@Ludwig.2019], I use 4 and 8
- low general reflectance

# 

# References
